Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE)

### 2014-02-24 WSSSPE1_report/WSSSPE1_summary_report.tex
A few more paragraphs from the questions
This will move into more cohesive sections soon!

>>>>

Are there significant differences in projects that have been running for 1, 3,
5, or 10+ years? Are there shared experiences for projects of a similar stage
of maturity? It was noted that computing and communication have changed
significantly over the past decade, and many of the experiences are tied to
the history of computing and communication. See the history of GCC, Emacs, or
the Visualization Toolkit for examples. Others felt that computing has changed
less, but communication and the widespread availability of tools has. It was
noted that email lists, websites, chat rooms, version control, virtual and
physical meetings are all over 20 years old.

There was debate that while some of the basics of computing may be fairly
similar, that the tools commonly used for computing have changed quite
significantly. Reference was made to Perl, which was commonly used, giving way
to whole new languages, such as Python, for gluing things together and how this
induces many students into entirely rewriting the scaffolding, leaving the old
to rot and the experiments to become non-reproducible as the tools change.
Jason Riedy stated that he had been guilty of this in the past. There was
discussion of this tendency along with the enormous differences in the speed and
ease of sharing---having to ship tapes around (GCC, Emacs, etc) as opposed to
the immediate sharing of the latest development in CVS, Subversion, Git,
Mercurial, Bazaar, etc.

The question was also posed as to whether the distinction between researcher
and developer is sensible, with James Hetherington commenting that in the UK
they have been examining a more nuanced view of research software engineers
and researcher developers. Should this be less of a contract relationship, and
more of a collaborative relationship. This is also at the core of the business
model that Kitware presented in its submission to the workshop. Are other
ingredients missing such as applied mathematicians? Should this be defined more
in terms of skill sets rather than roles and/or identities? This builds on the
comments from Vaidy Sunderam that scientists are generally good writers, and
have mathematical skills, so why can't they learn software engineering
principles?

Miller commented that all of the infrastructure that sits around a new
algorithm that we need to make it useful and sustainable requires different
skill sets than the algorithm developer. Friere commented that there are no
good career paths for people with broad skills, no incentives for them to
continue in these roles. There was debate around people doing what interests
them, and learning computing leaves people cold, but is it that it leaves the
people who find career paths in academia cold versus the full spectrum of
people involved in research? Is this also caused by poor teaching, or that the
perceived benefits for doing this are too small? It could also be attributed to
their focus being on science, not software engineering, or do the people with
the passion for software engineering in science simply have no viable career
path and either adapt or seek out alternate career paths.
